---
title: "Assignment B4"
author: "Katia Rosenflanz"
date: "2024-12-03"
output: github_document
---
## Overview 
For Assignment B4, Option A- Strings and Functional Programming in R was selected. 

Solutions to both **Exercise 1** and **Exercise 3** are given below. 

### Exercise 1

```{r}
library(janeaustenr)
library(tidyverse)
library(tidytext)
library(palmerpenguins)
library(broom)
```


For this exercise, the book *Persuasion* from the `janeaustenr` package will be chosen for analysis.
The `tidytext` package is used to define "stop words" in the book. 
We can extract all words that are **not** included in tidytext::stop_words with the help of the `tidytext` package, which has very useful functions for text mining. 

```{r}
words <- tibble(text = persuasion) %>% #convert "Persuasion" into tibble  
  #unnest_tokens() converts to lowercase, removes punctuation, and splits into one word per row
  unnest_tokens(output = word, input = text) %>%
  anti_join(stop_words, join_by(word)) %>% #remove all rows with a match in stop_words
  count(word, sort = TRUE) #counts unique words and sorts by most frequent

```

We can now check how many unique words there are in *Persuasion* (excluding stop words) by checking the length of the "word" column.  
```{r}
length(words$word)
```
There are 5340 unique words in the book (excluding stop words). It would be impractical to plot all of the words, therefore we will plot only the first 25. 
```{r}
#extract top 25 most frequent words 
top25 <- words[1:25,]
#plot word frequency
ggplot(top25, aes(x = reorder(word, n), n)) + 
  geom_col(fill = "lightblue") + 
  coord_flip()+ #move words to y axis 
  labs(title = expression(paste("Top 25 Most Frequent Words in ", italic("Persuasion"))),
                                x = "Words", y = "Frequency") #create title with italicized book title

```

### Exercise 3

For the following data analysis, I will be utilizing the **penguins** dataset from the `palmerpenguins` package. I will assess how the flipper length (mm) is related to the body mass (g) of a penguin. I will do so using a linear regression model.
The model will be fit to each species of penguin (Adelie, Gentoo, and Chinstrap) separately, and will be used to generate "predicted" flipper length values for each observed body mass value. 

```{r}
#create linear model using penguins dataset 
penguin_model <- penguins %>%
  #select only relevant columns
  select(species, body_mass_g, flipper_length_mm) %>%
  #nest data by species
  nest(data = c(body_mass_g, flipper_length_mm)) %>%
  #fit linear model to each species (making column of model objects)
  mutate(model = map(data, ~lm(flipper_length_mm ~body_mass_g, data = .)))

#print intermediate tibble for inspection 
penguin_model
```
The intermediate tibble above shows that both the data and model columns have nested tibbles for each species of penguin. 

Next, the `broom` package can enable us to extract further information about the linear models, specifically about the predicted values generated using the model. 
This means that we can take a look at what flipper length the model predicts based on a given observation of penguin body mass. 

```{r}
#using broom::augment, can extract the fitted values for predicted flipper length based on each observed body mass value
#and the standard error of the predicted values
penguin_results <- penguin_model %>%
  mutate(species, yhat = map(model, ~augment(.x, se_fit = TRUE)), .keep = "none") %>%  #adds information about each observation using the model
  unnest(yhat) %>% #each data point has its own row, unnested
  select(species, flipper_length_mm, body_mass_g, .fitted, .resid, .se.fit) %>% #select only useful columns
  rename(predicted_flipper_length = .fitted, residual = .resid, se_fit = .se.fit) #rename columns for readability

#print cleaned unnested tibble 
penguin_results

```
We can next calculate 95% confidence intervals for the predicted values, using the standard errors of the fit. This gives us an idea whether the predictions have low or high uncertainty based on the model.

```{r}
final_penguin <- penguin_results %>%
  mutate(low_ci = predicted_flipper_length -1.96*se_fit) %>% #lower bound of 95% confidence interval
  mutate(high_ci = predicted_flipper_length + 1.96*se_fit) %>% #higher bound of 95% confidence interval
  select(species, flipper_length_mm, body_mass_g, predicted_flipper_length, low_ci, high_ci)

#print the final tibble
final_penguin
```

Finally, we can generate a graph showing the observed data points, as well as the predicted data points from the linear model for each of the three species of penguin. The 95% confidence interval for the predicted values is also shown as a shaded area around the fits.

```{r}
# Visualization of model fit for flipper length versus body mass, colored by species of penguin
ggplot(final_penguin, aes(x = body_mass_g, y = flipper_length_mm, color = species)) +
  geom_point(alpha = 0.5) +
  geom_line(aes(y = predicted_flipper_length), linewidth = 1) + #show predicted values
  geom_ribbon(aes( ymin = low_ci, ymax = high_ci, fill = species), alpha = 0.3) +  #add 95% confidence interval
  labs(title = "Penguin Flipper Length vs Body Mass by Species",
    y = "Flipper Length (mm)",
    x = "Body Mass (g)") +
  theme_bw()
```

For each of the three species of penguin, there is a positive correlation between the body mass and flipper length based on the linear model. Most Gentoo penguins demonstrate larger body mass and longer flipper length compared to the other two species of penguin. The slope of the curve is also steeper for the Chinstrap penguin; this indicates that for small increases in body mass, flipper length increases more for the Chinstrap compared to the Adelie. Each of the shaded confidence intervals are fairly narrow, indicating that the standard errors of the fit were quite low and that the model's uncertainty is fairly low in predicting flipper length.  
